Cognitive and ecological aspects determine cleaner performance

### 2021-12-07 manuscript_1.0.Rmd
New file to make figures for grouped models. Updated source code to allow for predictions with grouped data.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Zeni_inpress]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load(Grutter xxxx). However, early laboratory
experiments showed that leaving causes service priority independent of
client value [@bshary_Asymmetric_2002]. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014].

@@

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
@dubois_Model_2021). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task [@prat_Role_2021;
@quinones_Reinforcement_2019]. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
[@quinones_Reinforcement_2019]: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e. as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning (Thorndike 1917), this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

@@

In order to capture with the model the relationship between the
characteristics of the market (environment) and the cleaner's
performance we need to translate the absolute abundances of the data to
relative abundances. This is because in the model, relative abundances
of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g. there are no
clients to be cleaned). The more cleaners there are, the more options
are available for the clients and hence the higher is the probability of
empty cleaning stations. How often the cleaning stations are empty in
turns influences the value that cleaners can expect from their choices.
The emptier the cleaning stations are, the more benefitial it is to
prioritize a client that will not stay. However, the frequency of empty
spots in the cleaning station does not only depend on cleaner relative
abundances but also on the frequency with which clients visit the
station. The client visitation frequency will likely vary among
different clients depending on parasite abundances and loads. These are
variables for which we do not have field estimates. Thus, we computed a
measure of relative cleaner abundance for each site relative to absolute
abundances and multiplied by a scaling constant that changes the range
of the variable. We fit the value of the scaling constant as part of the
statistical inference. As for the client abundances for residents and
visitors we computed a relative measure with respect to the total client
abundance, and weighted that by the rescaled relative absence of
cleaners (1-relative cleaner abundance). Thus, all three measures of
relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.

>>>>

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011,@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995a]. However, 
early laboratory
experiments showed that leaving causes service priority independent of
client value [@bshary_Asymmetric_2002]. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014].

@@

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
@dubois_Model_2021). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task [@prat_Role_2021;
@quinones_Reinforcement_2019]. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
[@quinones_Reinforcement_2019]: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e. as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

@@

In order to capture with the model the relationship between the
characteristics of the market (environment) and the cleaner's
performance we need to translate the absolute abundances of the data to
to  a measure of relative abundances that captures client visitation patterns.
This is because in the model, relative abundances
of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g. there are no
clients to be cleaned). The more cleaners there are, the more options
are available for the clients and hence the higher is the probability of
empty cleaning stations. How often the cleaning stations are empty in
turns influences the value that cleaners can expect from their choices.
The emptier the cleaning stations are, the more beneficial it is to
prioritize a client that will not stay. However, the frequency of empty
spots in the cleaning station does not only depend on cleaner abundances 
but also on the frequency with which clients visit the
station. The client visitation frequency will likely vary among
different clients depending on parasite abundances and loads. These are
variables for which we do not have field estimates. Thus, we computed a
measure of relative cleaner abundance for each site relative to absolute
abundances and multiplied by a scaling constant that changes the range
of the variable. This scaling constant is meant to capture variation in the market 
conditions driven partly by cleaner abundance. We fit the value of the 
scaling constant as part of the
statistical inference. As for the client abundances for residents and
visitors we computed a relative measure with respect to the total client
abundance, and weighted that by the rescaled relative absence of
cleaners (1-relative cleaner abundance). Thus, all three measures of
relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.
